Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-10-26 inference.tex
Encoding to GKS.

\noindent {\bf Example} Let us reconsider the  TPC-C \C{new\_order}
transaction from Sec.~\ref{sec:motivation}. Recall that the state
transformer for the \C{foreach} loop shown in
Fig.~\ref{fig:foreach_code} is (\C{k1}, \C{k2}, and \C{k3} are
constants):
\begin{smathpar}
\begin{array}{l}
  \lambda(\stl,\stg).~ \stl \cup \C{item\_reqs}\bind
      (\lambda\C{item\_req}.~ \F_U(\stl,\stg) \cup \F_I(\stl,\stg))
\end{array}
\end{smathpar}
where:
\begin{smathpar}
  \begin{array}{lcl}
    \F_U & = & \lambda(\stl,\stg).~ \stg \bind(\lambda s. 
                \itel{{\sf table}(s) = \C{Stock} \conj 
                  s.\C{s\_i\_id} = \C{item\_req.ol\_i\_id}\\
         &   & \hspace*{0.9in}}
                     {\{ \langle \C{s\_i\_id}=s.\C{s\_i\_id};\, 
                                 \C{s\_d\_id}=s.\C{s\_d\_id};\,
                                 \C{s\_qty} = \C{k1}\rangle \}\\
         &   & \hspace*{0.9in}} {\emptyset}) \\
    \F_I & = & \lambda(\stl,\stg).~ \{\langle\C{ol\_o\_id}=\C{k2},\,
                 \C{ol\_d\_id}=\C{k3},\, \C{ol\_i\_id}=\C{item\_req.ol\_i\_id},\, 
                 \C{ol\_qty}=\C{item\_req.ol\_qty}\rangle\}\\
  \end{array}
\end{smathpar}
For any $\stl$ and $\stg$, $\F_U(\stl,\stg)$ and $\F_I(\stl,\stg)$ are
expressions in $\SL$, so can be translated to FOL by the encoding
algorithm in Fig.~\ref{fig:logic}. Since the iteration variable
\C{item\_req} occurs free in these expressions, the appropriate
application of the encoding algorithm is
$\mssemof{\C{item\_req}}{\F_U(\stl,\stg)}$ and
$\mssemof{\C{item\_req}}{\F_I(\stl,\stg)}$, which results in
$(\phi_U,\G_U)$ and $(\phi_I,\G_I)$, respectively, where $\phi_U$,
$\phi_I$, $\G_U$, $\G_I$ are as shown below:
\begin{smathpar}
  \begin{array}{lcl}
    \phi_U & = & \forall \C{item\_req}.\forall s. \forall s'.~  \\
    & & \hspace*{0.1in}\pi(\C{item\_req}) \Leftrightarrow(s \in \stg) \conj 
        (\itec{{\sf table}(s)=\C{Stock} \wedge 
              s.\C{s\_i\_id} = \C{item\_req.ol\_i\_id}\\
    & & \hspace*{1.45in}}
              {s' = \langle \C{s\_i\_id}=s.\C{s\_i\_id};\, 
                            \C{s\_d\_id}=s.\C{s\_d\_id};\,
                            \C{s\_qty} = \C{k1}\rangle\\
    & & \hspace*{1.45in}} {\bot}) \\
    & & \hspace*{1.2in} \Leftrightarrow g(\C{item\_req},s')\\
    \G_U & = & \lambda (\C{item\_req},r).~ \pi(\C{item\_req}) 
        \wedge g(\C{item\_req},r)\\
    \phi_I & = & \top \\ 
    \G_U & = & \lambda (\C{item\_req},r).~ 
        r = \langle \C{ol\_o\_id}=\C{k2},\,
                    \C{ol\_d\_id}=\C{k3},\, 
                    \C{ol\_i\_id}=\C{item\_req.ol\_i\_id},\, 
                    \C{ol\_qty}=\C{item\_req.ol\_qty}
             \rangle\\
  \end{array}
\end{smathpar}

>>>>

\noindent {\bf Example} Let us reconsider the  TPC-C \C{new\_order}
transaction from Sec.~\ref{sec:motivation}. Recall that the state
transformer ($\T$) for the \C{foreach} loop shown in
Fig.~\ref{fig:foreach_code} is (\C{k1}, \C{k2}, and \C{k3} are
constants):
\begin{array}{l}
  \lambda(\stl,\stg).~ \stl \cup \C{item\_reqs}\bind
      (\lambda\C{item\_req}.~ \F_U(\stl,\stg) \cup \F_I(\stl,\stg))
where:
    \F_U & = & \lambda(\stl,\stg).~ \stg \bind(\lambda s. 
                \itel{{\sf table}(s) = \C{Stock} \conj 
                  s.\C{s\_i\_id} = \C{item\_req.ol\_i\_id}\\
         &   & \hspace*{0.9in}}
                     {\{ \langle \C{s\_i\_id}=s.\C{s\_i\_id};\, 
                                 \C{s\_d\_id}=s.\C{s\_d\_id};\,
                                 \C{s\_qty} = \C{k1}\rangle \}\\
         &   & \hspace*{0.9in}} {\emptyset}) \\
    \F_I & = & \lambda(\stl,\stg).~ \{\langle\C{ol\_o\_id}=\C{k2},\,
                 \C{ol\_d\_id}=\C{k3},\, \C{ol\_i\_id}=\C{item\_req.ol\_i\_id},\, 
                 \C{ol\_qty}=\C{item\_req.ol\_qty}\rangle\}\\
For any $\stl$ and $\stg$, $\F_U(\stl,\stg)$ and $\F_I(\stl,\stg)$ are
expressions in $\SL$, so can be translated to FOL by the encoding
algorithm in Fig.~\ref{fig:logic}. Since the iteration variable
\C{item\_req} occurs free in these expressions, the appropriate
application of the encoding algorithm is
$\mssemof{\C{item\_req}}{\F_U(\stl,\stg)}$ and
$\mssemof{\C{item\_req}}{\F_I(\stl,\stg)}$, which results in
$(\phi_U,\G_U)$ and $(\phi_I,\G_I)$, respectively, where $\phi_U$,
$\phi_I$, $\G_U$, $\G_I$ are as shown below:
    \phi_U & = & \hspace*{0.2in}\forall \C{item\_req}.\forall s. \forall s'.~ 
        \pi_1(\C{item\_req}) \Leftrightarrow \\
    & & \hspace*{0.3in}(s \in \stg) \conj 
        (\itec{{\sf table}(s)=\C{Stock} \wedge 
              s.\C{s\_i\_id} = \C{item\_req.ol\_i\_id}\\
    & & \hspace*{1.45in}}
              {s' = \langle \C{s\_i\_id}=s.\C{s\_i\_id};\, 
                            \C{s\_d\_id}=s.\C{s\_d\_id};\,
                            \C{s\_qty} = \C{k1}\rangle\\
    & & \hspace*{1.45in}} {\bot}) \Rightarrow g_0(\C{item\_req},s')\\
    & & \conj \hspace*{0.08in}\forall \C{item\_req}.\forall s'. \exists s.~ 
        \pi_2(\C{item\_req}) \Leftrightarrow \\
    & & \hspace*{0.28in} g_0(\C{item\_req},s') \Rightarrow (s \in \stg) \conj
        (\itec{{\sf table}(s)=\C{Stock} \wedge 
              s.\C{s\_i\_id} = \C{item\_req.ol\_i\_id}\\
    & & \hspace*{1.8in}}
              {s' = \langle \C{s\_i\_id}=s.\C{s\_i\_id};\, 
                            \C{s\_d\_id}=s.\C{s\_d\_id};\,
                            \C{s\_qty} = \C{k1}\rangle\\
    & & \hspace*{1.8in}} {\bot}) \\
    \G_U & = & \lambda (\C{item\_req},r).~ \pi_1(\C{item\_req}) \wedge 
        \pi_2(\C{item\_req}) \wedge g_0(\C{item\_req},r)\\
        \phi_I & = & \top \\ 
    \G_I & = & \lambda (\C{item\_req},r).~ 
        r = \langle \C{ol\_o\_id}=\C{k2},\,
                    \C{ol\_d\_id}=\C{k3},\, 
                    \C{ol\_i\_id}=\C{item\_req.ol\_i\_id},\, 
                    \C{ol\_qty}=\C{item\_req.ol\_qty}
             \rangle\\
Since the transformer ($\T$) of the \C{foreach} loop is not nested
does not contain any free iteration variables, the appropriate
application of the encoding algorithm  is
$\mssemof{\emptyset}{\T(\stl,\stg)}$, which results in the
$(\phi,\G)$, as defined below:
    \phi & = & \phi_I \conj \phi_U \conj \\
    & & \hspace*{0.2in}\forall \C{ireq}. \forall s.~ 
        \pi_3 \Leftrightarrow (\C{ireq} \in \C{item\_reqs} \conj 
        (\G_U(\C{ireq},s') \vee \G_I(\C{ireq},s'))
        \Rightarrow g_1(s))\\
    & & \conj \hspace*{0.05in}\forall s.\exists \C{ireq}. ~ 
        \pi_4 \Leftrightarrow (g_1(s) \Rightarrow (\C{ireq} \in \C{item\_reqs} 
        \conj (\G_U(\C{ireq},s') \vee \G_I(\C{ireq},s')))\\
    \G & = & \lambda(r).~\pi_3 \wedge \pi_4 \wedge g_1(r)\\
